### 2009-12-02 Paper/paper.tex
Paper update

\documentclass[letterpaper]{article}
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{url}
\usepackage[table]{xcolor}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{float}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PDFMARK for TeX and GhostScript
% Uncomment and complete the following for metadata if
% your paper is typeset using TeX and GhostScript (e.g
% if you use .ps or .eps files in your paper):
% \special{! /pdfmark where
% {pop} {userdict /pdfmark /cleartomark load put} ifelse
% [ /Author (John Doe, Jane Doe)
% /Title (Paper Title)
% /Keywords (AAAI, artificial intelligence)
% /DOCINFO pdfmark}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PDFINFO for PDFTeX
% Uncomment and complete the following for metadata if
% your paper is typeset using PDFTeX
\pdfinfo{
  /Title (Solving an Exam Scheduling Problem Using a Genetic Algorithm)
  /Subject (Input The Proceedings Title Here)
  /Author (Dave, Kordalewski;
           Caigu, Liu;
           Kevin, Salvesen;)
} 
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Uncomment only if you need to use section numbers
% and change the 0 to a 1 or 2
% \setcounter{secnumdepth}{0}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Solving an Exam Scheduling Problem Using a Genetic Algorithm}
\author{Dave Kordalewski \\ davekordalewski@gmail.com
   \And Caigu Liu \\liucaigu@gmail.com
   \And Kevin Salvesen \\ kevin.salvesen@gmail.com}

\nocopyright

\begin{document}
\maketitle

\begin{abstract}
  We examine the exam scheduling problem with soft constraints for problems of varying
  sizes using a genetic algorithm (GA), random search, hillclimbing, and a variant of simulated
  annealing. This is done in the context of maximizing the fitness of a schedule, where the
  fitness is relatively computationally expensive to compute for any schedule. The genetic
  algorithm is discussed in detail. We consider selecting parameters for the GA that
  maximize the quality of the schedule found after a constant number of fitness computations. 
  The GA, with a naive crossover operator, performs much better than the random search, but
  simulated annealing is far superior in this search space.
\end{abstract}

\section{The Problem}
  The problem we investigate is a sort of scheduling problem. 
  We are attempting to find the most satisfactory choice of when
  and where to hold exams, given a background of student course loads.

  An instance of this scheduling problem consists of a number of days
  on which exams can be scheduled ($d$), the number of time slots in which
  an exam can be scheduled on any day ($t$), a set of rooms ($R$), a set of
  courses ($C$), and a set of students ($S$), each of whom ($s$) has a
  particular course load, some subset of the courses ($L_s$).

  A schedule, then, is a mapping from courses to rooms at times, 
  which we can express like this:

  \[ C \rightarrow \left\{(r,\ a,\ b)\ |\ r\in R,\ a\in \{1..d\},\ b\in \{1..t\}\right\} \]

  The total number of different possible schedules in any problem instance is
  $|C|^{|R|dt}$ which is typically far too large for brute force search.
  For example, one instance that we examine (scheduling the fall semester
  exams at the University of Toronto in 2009) involves 603 courses, 7 days,
  8 times, 43 rooms, and 21945 students. This instance admits approximately 
  $10^{6695}$ different possible schedules.

  Typically, with relatively loose constraints on the number of rooms and 
  particular schedules of students, it is not difficult to find some consistent
  schedule; that is, one where no student is asked to write 2 exams
  simultaneously and no room has 2 exams occur in it simultaneously.

  Rather than worry about hard constraints, we prefer a framework of soft 
  constraints, where we try to find a schedule that makes the students and 
  invigilators happiest overall, allowing the possibility that some student 
  or room is left with an impossible exam schedule, which can, in the real 
  world, be dealt with on an individual basis.

  The timetable that a student $s \in S$ has under some particular schedule 
  $K$ may be represented in this way:

  \[ TT_s(K) = \left\{(a, b)\ |\ \exists r \in R, \exists c \in L_s, K(c)=(r, a, b) \right\} \]

  which may, in general, be a multiset.

  Similarly, the timetable for a room $r \in R$ is

  \[ TT_r(K)=\left\{(a, b)\ |\ \exists c\in C,\ K(c)=(r, a, b)\right\} \]

  We define two quality functions, mapping schedules to real numbers in [0,1],
  one for students and one for rooms. These are meant to capture how much they
  "like" the schedule under consideration. (For instance, a student will rate
  poorly any schedule where she has 2 consecutive exams, and very poorly any 
  schedule which expects her to take two exams at the same time. A room will 
  rate poorly a schedule that assigns two different exams to be held in it at 
  the same moment).

  \[ Q_s(K)=q_{student}\left(TT_s(K)\right):\textit{ Schedules}\rightarrow [0,\ 1] \]
  \[ Q_r(K)=q_{room}\left(TT_r(K)\right):\textit{ Schedules}\rightarrow [0,\ 1] \]

  We will consider later how to define these functions in a reasonable way.

  The quality of a schedule, then, is given by

  \[ Q(K)= \frac{\left(w_s\left(\sum_{s \in S}{\frac{Q_s(K)}{|S|}}\right) +
     w_r\left(\sum_{r\in R}{\frac{Q_r(K)}{|R|}} \right)\right)}{(w_s+w_r)} \]

  where $w_s$ and $w_r$ are the weightings we can use to indicate that we 
  care somewhat more about students preferences than rooms, giving a quality
  (or fitness, in the terminology of genetic algorithms) in the range [0,1] for
  any schedule.

  Evaluating this function Q can be computationally expensive when $|S|$ and $|R|$
  are large. Our work examines how to find a schedule with relatively high fitness
  considering that we will want to do this with as few evaluations of Q as possible.

  We examine, first and foremost, a genetic algorithm approach to solving this problem,
  but also touch on a few other methods.

\section{Approach and Implementation}
  \subsection{Differences From Natural Exam Scheduling Problems}
    The problem we have constructed here is a compromise between the most
    mathematically pure problem with the same sort of properties (which might ignore the
    rooms and related constraints completely) and the problem that must be solved in the
    real world when scheduling thousands of students, which has some important differences
    from this problem. For example, we don't deal with constraints such as room capacity
    or exams with varying length.

  \subsection{Tools used}
    Our application was entirely written in Java. We also wrote another Java application 
    to generate problem instance files for testing our algorithms. We used Mathematica 
    to plot all of our graphs.  Additionally, we use Google code's source code
    repository\footnote{\url{http://code.google.com/p/csc384-genetic-algorithms-project/}} 
    and Google docs to collaborate on this report.

  \subsection{Fitness Function}
    In order to assign each schedule a fitness, we need to define the functions 
    $q_{student}$ and $q_{room}$. They need to capture the idea of determining how
    much each student or room likes their given timetable. We do this by calculating 
    the penalty incurred by a room, which is a measure of it having properties that 
    are unpleasant to the student or room, and then using the formula

    \[ q(TT)=\frac{1}{1+penalty_{TT}} \]

    This function will be in the desired range or [0,1].

    For students, we accrue a penalty of 50 for every time that is repeated in a 
    timetable, 5 when there are exams in 2 consecutive times, 3 when 2 exams happen 
    on the same day, and 0.5 when the student has exams on consecutive days. This 
    accords reasonably well with real students' hopes for their exam timetables.

    Rooms have the same penalty of 50 when it is used by two exams at the same time, 
    5 if it hosts exams in consecutive time slots (reasoning that the invigilators 
    need time to let the students out and in and prepare for the next exam) and a 
    penalty of 0.5 when a room is empty for an entire day between uses (reasoning 
    that the room requires some amount of modifications to be used for exams, and 
    it is difficult to use it for other purposes if it will still be used for exams 
    soon.)

    We chose the weighting constants $w_s=2$ and $w_r=1$, which suggest that we are 
    more concerned with the students' schedules than the rooms'.

    The details of the function we used are essentially arbitrary, and we expect 
    similar results would be attained for any other penalty-based fitness function 
    modelling the preferences of students and requirements of rooms.

  \subsection{Genetic Algorithm Search}
    Although there are many readily available open source genetic algorithm software
    packages available (see JGAP, Jenes) we decided to write a simple, yet general, 
    genetic algorithm package, in Java, that can interface with data types designed 
    by others, so long as they implement a few necessary methods. The package we wrote 
    is more than adequate for investigating the scheduling problem under consideration here.

    \subsubsection{General Description}
      The genetic algorithm maintains a population of schedules and constructs subsequent
      generations by adding new random schedules, making copies of high quality schedules 
      in the present generation, applies the mutation operator to schedules, or uses the
      crossover operator on 2 schedules. When performing mutation or crossover, the 
      schedules are selected with a probability proportional to their fitness, allowing 
      the schedules with high quality to be selected more frequently and make more 
      ``offspring''.

      The free parameters in our GA are the size of the population, number of
      generations to simulate, and proportions of new generations to be made by the copy, 
      random, mutation, and crossover methods.

    \subsubsection{Representation}
      We use the naive representation of schedules. A schedule contains lists of size $|C|$
      containing the room in which each course's exam is to occur and the timing
      (corresponding to a pair or integers representing the day and time) of each course's
      exam.

      We will see that this representation is far from ideal, but there is no clear 
      alternative.

    \subsubsection{Mutation Operator}
      The mutation operator takes a probability and a schedule. We loop over the courses,
      and, with the given probability, each is or is not set to a random room at a random time.

    \subsubsection{Crossover Operator}
      Given two schedules, the crossover operator loops over the courses, and chooses at
      random which schedule to take the room and timing from for that course. This results 
      in an interleaving of the original courses' data.

      There is, unfortunately, no good reason to expect that the result of applying 
      crossover to two relatively high quality schedules will produce a high quality 
      schedule, since the quality of schedules can be very subtle, as they are a more 
      global property of the schedule than this process captures. The crossover of two 
      schedules where no student or room has a timing conflict is unlikely to result in 
      a schedule with much more than average quality. There is no obvious reason to think 
      that this sort of procedure will produce schedules with better than random fitness. 
      Luckily, the schedules produced in practice are better than random.

      If a different representation were chosen, such that the children schedules were
      different from, but likely to have quality similar to their parents, the GA would 
      likely be able to perform much better than it does, but we could not discover such a
      representation and are satisfied with this method for present purposes.

  \subsection{Other Search Algorithms}
    \subsubsection{Random Search}
      The random search method simply generates random schedules, evaluates their fitness,
      and remembers the one that it has seen that has highest fitness. This method is meant
      as a baseline against which to compare the other methods. If a search method doesn't
      perform significantly better than random search with the same number of fitness
      evaluations, it is useless.

    \subsubsection{Hillclimbing Search}
      The hillclimbing search that we have implemented examines every schedule in the
      neighbourhood of a given starting schedule, and then continues from the one with
      highest fitness, terminating when it has performed as many evaluations as it is allowed
      to, or it has found a schedule whose neighbour are all equal or inferior to it.

      The neighbouring schedules are those where only one course meets in either a different
      room, on a different day, or at a different time. This leads to a total of
      $ |C|(d+t+|R|-3) $ schedules in the neighbourhood of any given schedule.

      It can be seen that this method is completely deterministic given a starting schedule,
      and will always return a local maximum

    \subsubsection{Mutate Search}
      Mutate Search starts by generating a random schedule. It makes an alternative schedule
      (by using the same mutate operator made for the genetic algorithm) with 1\% of the
      data randomized and chooses which of the schedules is better. This operation is
      repeated until it has performed as many evaluations as it is allowed to.

      This method can be changed to true simulated annealing by varying the degree of
      modification made to the current best schedule with the number of evaluations
      remaining. Not being the focus of our investigation, little effort was made to 
      optimize the parameters. 1\% does seem to give better performance than other nearby 
      values, though we did not collect data to substantiate this. Finding ideal parameters 
      for simulated annealing on this problem would make for a project-sized inquiry in itself.

    \subsection{Scheduling Problem Instances}
      We generated 2 sets of 3 scheduling problem instances with which to test the 
      above algorithms. The first set of instances were generated with the parameters 
      in the first three columns of the table below. The rooms, courses and students 
      were given arbitrary names and the courses that each student attended were 
      selected with uniform probability from the complete set of courses.

      The second set was generated with a sort of correlation between the courses that each 
      student took, rendering the dataset more true to the sort of data that would be found 
      in real-world situations. The set of courses that each student takes was selected such 
      that we could expect that 70\% of their courses were courses in their major. This 
      simulates the way in which we expect student bodies to choose their courses, though 
      the true data is likely to exhibit far more structure, due to program requirements, 
      notorious professors, etc.

      The 3 instances' size vary from manageable to large. The smallest could potentially be 
      solved exactly, with methods not under consideration here, while the others have 
      search spaces much too large for any hope of finding a perfect solution. For the 
      largest, which simulates a UofT exam schedule, a single evaluation of the fitness 
      function takes 0.2 seconds on on a typical desktop computer. This makes it difficult 
      to thoroughly investigate this instance.

      The following table characterizes the problem instances that we used ($\#$Maj is the 
      number of different majors used and \textsc{SSS} is an approximation of the search space size):    
      \definecolor{light-gray}{gray}{0.87}
      \rowcolors{1}{light-gray}{white}
      \begin{center}
      \begin{small}
      \begin{tabular}{*{7}{c}}
      \hline 
      	& S 	& M 	& L 	& $S_{Cor}$ & $M_{Cor}$ & $L_{Cor}$ \\ 
      \hline \hline
       d 		& 7 	& 10	& 7 	& 7 	& 10 	& 7 	\\ 
       t 		& 5		& 8 	& 8 	& 5 	& 8 	& 8 	\\ 
       $|C|$	& 20 	& 200 	& 603 	& 20 	& 200 	& 600 	\\ 
       $|R|$	& 4 	& 10 	& 43 	& 4 	& 10 	& 43	\\ 
       $|S|$	& 50 	& 300 	& 21945 & 50 	& 300 	& 21945	\\ 
       $|L_s|$	& 1-5 	& 1-6 	& 1-5 	& 2-6 	& 2-6 	& 2-6 	\\ 
       \begin{tiny}$\#$Maj\end{tiny}	& N/A 	& N/A 	& N/A 	& 4 	& 10 	& 20 	\\ 
       \textsc{SSS}	& $10^{182}$ & $10^{1840}$ & $10^{6695}$ & $10^{182}$ & $10^{1840}$ & $10^{6689}$ \\
      \hline 
	  \end{tabular} 
	  \end{small}
	  \end{center}

	  The large data set was created with parameters inspired from the University of 
	  Toronto's Faculty of Arts and Science.

\section{Evaluation}
   \subsection{Selecting Parameters for the GA}
      In order to select parameters for the genetic algorithm (that is, the population, 
      number of generations to simulate, proportions of copies, randoms, mutations, and 
      crossovers for the next generation) we performed tests using the small uncorrelated 
      problem instance.

      \paragraph{Generation Parameters}
        We undertook a search of the space of parameters for forming new generations.

        Our results can be seen on figure 1, where each data point of the graph 
        represents the average of 5 trials. To make the scale clearer, the colour 
        on the surface corresponds to the fitness, where the highest fitnesses 
        are red. The number of copies is fixed at 1, and the proportions of 
        mutations and crossovers are shown on the horizontal axes.

        \begin{figure}[h]
  		  \centering
  		  \includegraphics[width=8cm]{GAParametersData}
  	      \caption{Effectiveness of the GA's proportion parameters}
  	      \label{graph:GAParam1}
		\end{figure}

        The graph indicates that for a fixed number of evaluations, best results are 
        achieved by selecting parameters that have the majority of individuals in the next 
        generation be generated by crossovers of the most fit schedules in the previous 
        generation. In order to achieve the very highest fitness, a few randoms or mutations 
        should also be incorporated.

        This test was done for only the smallest problem instance, and is infeasible for the 
        larger instances. We chose generation proportion parameters to approximate the optimal 
        values seen here, under the assumption that the larger problems exhibited this sort of 
        preference.

      \paragraph{Size Parameters}
        We also performed tests to gauge the optimal population size and number of generations 
        to calculate. We determined that very small values for either the population or number 
        of generations result in very poor generated solutions. A population of around 50 is 
        sufficient for good diversity to be retained in the population, and having a number of 
        generations relatively large compared to the population size seemed to be effective.

        We gathered data from the small and medium problem instances, but the precise qualities 
        provided by particular parameters will not necessarily hold in larger instances. The 
        optimal population size and number of generations (for a fixed number of fitness 
        evaluations) is likely to vary for larger problem instances, but it seemed clear that 
        the number of generations should be set somewhat higher than the population.

  \subsection{Comparison Between the Different Searches}
    We tested the 4 search methods on all 6 of our problem instances. The results 
    for two of the problems, the small and large problems, are shown on figure 2 
    and 3. The medium problem exhibited characteristics very similar to the large problem.

    All methods produced nearly indistinguishable results for the uncorrelated 
    versions. This suggests that it may be possible to use the expected internal 
    correlations more effectively that the methods we have proposed do.

    On the graphs for the small instance, the results shown are averages over many trial 
    runs. Individually, however, the different methods all produced similar results, in 
    particular the ranking of the algorithms was very consistent for any particular problem 
    instance.

    \begin{figure}[h]
  	  \centering
  	  \includegraphics[width=8cm]{SMALL-TYPE2_FullAverageStats}
      \caption{Comparison of the different searches on a small data set}
	\end{figure}

	\begin{figure}[h]
  	  \centering
  	  \includegraphics[width=8cm]{HUGE-TYPE2_FullAverageStats}
      \caption{Comparison of the different searches on a large data set}
	\end{figure}

	Random search provides a benchmark by which to measure the other 3 techniques. 
	As desired, they all usually perform significantly better.

	Hillclimbing works very well for the small instance, but miserably on the large 
	problem. This is because of the vastly increased number of neighbouring schedules 
	in the large problem. Hillclimbing is able to examine a much smaller portion of 
	search space and wastes its evaluations on too many points that are too similar 
	to each other. A variant that tries some fixed number of schedules in the 
	neighbourhood may perform better.

	The mutate search algorithm performed remarkably well in all tests, finding a 
	schedule at least somewhat better than every other method for every size problem 
	instance.

	The GA is a bit of a disappointment. It works much better than random, meaning 
	that the naive crossover method used does have some value, and it scales well 
	for different sized problem instances, which is an important characteristic. 
	For the large instance, the attained fitness of around 0.4 would be the fitness 
	of a schedule for which every student and room assign 1.4 penalty points to the 
	schedule. This is reasonable, and would perhaps be very exciting if mutate search 
	wasn't so much better than GA for this problem.

\section{Conclusion}
  \subsection{Fitness Function Modifications}
    The fitness function can be modified by any monotone function to obtain identical 
    results for the 3 non-GA methods, which only need to choose the better fitness 
    and do not use the fitness values in any important way. The GA, however, uses the 
    fitness values to decide the probability of schedules being selected and modified 
    for subsequent generations. Modifying the fitness function in a variety of ways 
    (perhaps calculating the total penalty that a schedule gets instead averaging over 
    the values given by each student and room) might improve the quality of the GA 
    approach to this problem. The fitness function used here could be modified in many 
    ways, and investigating this would be very interesting.

\nocite{*}
\bibliographystyle{aaai}
\bibliography{paperBib}
\end{document}